---
title: "a3_task3_craig_kopulsky"
author: "Craig Kopulsky"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```


```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>%  # reading in urchins data
  clean_names() # converting variables to lowercase snake case
```


# **Section 1:** Size comparisons, purple and red urchins
Including data from all sites and for both treatments

## Exploratory data visualization
Comparing the distribution of sizes between purple and red urchins using histograms and box plots

```{r}
ggplot(data = urchins_data, aes(size)) + # creating histogram with the urchins data set showing distributions of sizes
  geom_histogram() +
  facet_wrap(~common_name) + # creating faceted histogram to show a comparison of size distributions between purple urchins and red urchins
  labs(x = "Urchin Diameter (cm), measured to the nearest .5cm",
       y = "Total Urchins") # updating labels
```

```{r}
ggplot(data = urchins_data, aes(x = common_name, y = size)) +
  geom_boxplot() + # creating box plot with urchins data set with the urchins dataset showing distribution of sizes
  labs(x = "Common Name", 
       y = "Urchin Diameter (cm), measured to the nearest .5cm") # labeling x and y axis
```

## Descriptive statistics table
Summary tables for purple and red urchins

```{r}
urchins_summary <- urchins_data %>% # creating summary table urchins_summary using urchins data set
  group_by(common_name) %>% # grouping by common_name
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n()) # adding means, medians, standard deviations, and sample sizes for each group to summary table
  kbl(urchins_summary) %>% 
  kable_styling(full_width = FALSE, "hover") # adding style to the summary table

```

## Summary of exploratory analysis
Red urchins appear to be larger than purple urchins on average, as the mean diameter for the red urchins sample (6.8 cm) is larger than the mean diameter for the purple urchins sample (4.5 cm). This mean diameter for the red urchins sample is about 51% larger than the mean diameter for the purple urchins sample. There is more variability in the red urchins sample as evidenced by the sd for that sample (1.9) compared to the sd for the purple urchins sample (1.2).

# **Section 2** Exploration of purple urchin counts by size and year

```{r}
# creating subset of urchins data focusing on purple urchins
purple_urchins <- urchins_data %>% # creating data set purple_urchins from urchins_data
  filter(common_name == "Purple Urchin") %>% # filtering by common_name to include only purple urchins
  mutate(date = mdy(date)) %>% # adding column for date in month/day/year format
  mutate(year = year(date)) %>% # adding column for year
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  )) # adding a column site_full which contains full site names
```

```{r}
# creating summary table containing total number of purple urchins observed by site and year
purple_urchins_summary <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarize(n = n())
```

```{r}
#creating exploratory graph of counts of purple urchins recorded at each site over time by year
ggplot(purple_urchins_summary, aes(x = year, y = n, color = site_full)) +
  geom_point()
```

