---
title: "a3_task3_craig_kopulsky"
author: "Craig Kopulsky"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```


```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>%  # reading in urchins data
  clean_names() # converting to lowercase snake case
```


# **Section 1:** Size comparisons, purple and red urchins
Including data from all sites and for both treatments

## Exploratory data visualization
Comparing the distribution of sizes between purple and red urchins

```{r}
ggplot(data = urchins_data, aes(size)) + # creating histogram with the urchins dataset showing distributions of sizes
  geom_histogram() +
  facet_wrap(~common_name) + # creating faceted histogram to show a comparison of size distributions between purple urchins and red urchins
  labs(x = "Urchin Diameter (cm), measured to the nearest .5cm",
       y = "Total Urchins") # updating labels
```

```{r}
ggplot(data = urchins_data, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name",
       y = "Urchin Diameter (cm), measured to the nearest .5cm")
```

## Descriptive statistics table
Summary tables for purple and red urchins

```{r}
urchins_summary <- urchins_data %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size), median = median(size), sd = sd(size), n = n())
  kbl(urchins_summary) %>% 
  kable_styling(full_width = FALSE, "hover")

```

## Summary of exploratory analysis
xxxxxxxxxx

# **Section 2** Exploration of purple urchin counts by size and year

```{r}
# creating subset of urchins data focusing on purple urchins
purple_urchins <- urchins_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```

```{r}
# creating summary table containing total number of purple urchins observed by site and year
purple_urchins_summary <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarize(n = n())
```

```{r}
#creating exploratory graph of counts of purple urchins recorded at each site over time by year
ggplot(purple_urchins_summary, aes(x = year, y = n, color = site_full)) +
  geom_point()
```

